---
title: "Final_Rep"
author: "Gauri Nagre"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





# Contribution of Each Group Member
#### Gauri Nagre
Role: Data Collection & Preprocessing,  Report Writing & Presentation
Contribution: Responsible for gathering the hospital dataset, cleaning missing data, created visualizations, drafted the report and ensuring data quality for analysis.

#### Rajwardhan Wagh
Role: Data Analysis & Visualization
Contribution: Conducted exploratory data analysis (EDA), created visualizations, and performed statistical analysis.

#### Akshita jaiswal
Role: Modelling & Interpretation
Contribution: Implemented regression models, analyzed results, and derived key insights.


____________________________________________________________________________________________________


# Introduction

### Overview of the Topic Area
Efficient hospital management is critical for ensuring patient satisfaction and operational success. Key challenges include high patient readmission rates, long waiting times, and suboptimal staff utilization. Data-driven solutions can help address these inefficiencies. The dataset contains **hospital data**, including **patient counts**, **readmission rates**, **waiting times**, and **staff utilization**. This aligns closely with the objectives outlined in the business proposal, which aims to enhance hospital efficiency and patient care through data-driven solutions


____________________________________________________________________________________________________



# Statement of the Business Problem
The hospital currently experiences:
•	A 47.2% readmission rate, indicating gaps in post-discharge care.
•	An average waiting time of 50 minutes, exceeding the target of 30 minutes.
•	Staff utilization inefficiencies that impact hospital workflow.


The goal is to reduce patient readmissions, decrease waiting times, and improve staff utilization efficiency through data-driven analysis.


____________________________________________________________________________________________________



# Methodology

#### Data Source and Variables
The dataset used for this analysis was sourced from [Hospital Database Name]. Key variables include:
•	Patient Readmission (Binary: Yes/No)
•	Waiting Time (Minutes)
•	Doctors/Nurses Attending (Count)
•	Staff Utilization (Workload Metric)
•	Medical Specialty

#### Data Preparation and Procedures
•	Data Cleaning: Removed missing values, handled outliers, and standardized variable formats.
•	Feature Engineering: Created derived variables (e.g., peak hours for patient visits).
•	Normalization: Applied where necessary to improve model performance.


#### Analytical Techniques Used
•	Exploratory Data Analysis (EDA): To identify trends and correlations.
•	Simple Linear Regression: To analyze the relationship between waiting time and staff availability.
Justification for Techniques:
•	Linear Regression is suitable for assessing continuous relationships (e.g., waiting time vs. nurses attending).
•	This method helps in understanding cause-and-effect relationships between two continuous variables.
•	It provides a quantifiable measure of how nurse availability affects waiting time, which directly ties into the business goal of reducing waiting times.


____________________________________________________________________________________________________



# Results

#### Findings from Analysis and Modelling
1. Model Performance Issues
Accuracy: 49.03% → Poor predictive power.

2. Significant Factors
Higher Readmission Risk:
Autoimmune Conditions (p = 0.0002).
Late-night & Afternoon Visits (p < 0.001).

Lower Readmission Risk:
Patients from Cardiology, Emergency, Oncology, and Surgery Wards (p < 0.05).

3. Model Instability
Some Length of Stay categories have extreme estimates → possible data imbalance.
Singularities (dropped variables) suggest collinearity issues.

4. Improvement
Tried tree-based models (Random Forest).
Removed highly correlated or insignificant variables but still gives the same acuracy



____________________________________________________________________________________________________



# Discussion

Reducing Readmissions (47.2%)
High-risk groups: Ages 13-35 & 60-75
Solutions: Scheduled nurse check-ins, medication reminders, collaboration with primary care
Decreasing Waiting Time (50 min, goal: 30 min)

More nurses = shorter wait times
Solutions: Staff redistribution, self-service check-in kiosks, smart triage system
Optimizing Staff Utilization (2,400 units inefficiency)

Solutions: Predictive analytics for staffing, cross-training, real-time monitoring system


____________________________________________________________________________________________________




# Recommendations

1. Length of Stay Impact
If any of the "Length of Stay" variables have a significant p-value (Pr(>|z|) < 0.05), shorter stays may reduce readmissions.
Recommendation: Optimize discharge planning to avoid unnecessarily long stays.

2. Age Group Influence
Significant coefficients in "age.(0-12)", "age.(13-35)", etc., indicate that certain age groups have higher/lower readmission risks.
Recommendation: Provide age-specific post-discharge care to reduce readmissions.

3. Medical Specialty & Ward Assignment
If "medical_specialty.Cardiology" or "Ward Name.Cardiology Ward" is significant, cardiology patients may need better post-discharge care.
Recommendation: Implement targeted follow-up plans based on specialty-specific risks.

4. Glucose Test & Diabetes Medication
"glucose_test" and "diabetes_med" significance suggests diabetes management affects readmission risk.
Recommendation: Improve glucose monitoring and medication adherence programs.

5. Hospital Visit Timing & Waiting Time
"Time Visited Hospital.1:00 AM - 5:00 AM", etc., could indicate readmission risk varies by visit time.
"Waiting Time" significance suggests high waiting times increase readmissions.
Recommendation: Optimize staffing and appointment scheduling to reduce wait times.



____________________________________________________________________________________________________




# Conclusion

Better staff allocation cuts wait times. Predictive analytics prevent readmissions. Smarter scheduling boosts efficiency.

Next Steps:
Use ML for real-time forecasting.
Expand data for broader insights.
Train staff with data-driven strategies.



____________________________________________________________________________________________________



# References
(dataset : https://www.kaggle.com/datasets/dubradave/hospital-readmissions
)


____________________________________________________________________________________________________




# LLM Use Declaration

We utilized LLMs in the following ways:
•	Brainstorming: Generating ideas for problem-solving approaches.
•	Code Reference: Assisting in writing and debugging R code.
•	Proofreading: Refining report structure and clarity.
•	Chatgpt : For code generation and correction
LLMs were not used for direct copy-pasting of content or automated decision-making. All final analyses and recommendations were reviewed and validated by the project team.




____________________________________________________________________________________________________



# Tests and Analysis:



```{r}
# Load necessary libraries
library(readxl)
library(caret)
library(dplyr)
library(ggplot2)

# Load dataset
data <- read_excel("C:/Users/Gauri Vijay Nagre/Desktop/Sem_2/BA/BA_Lifeline_Hospital_Dashboard/Hospital Dataset3.xlsx", sheet = "Sheet1")
head(data)
```

## Objectives:
#### 1. Reduce patient readmission rates by addressing underlying factors. 
#### 2. Optimize staff scheduling to align with peak patient visit times. 
#### 3. Shorten patient waiting times to improve satisfaction and care quality. 
#### 4. Enhance operational efficiency through targeted interventions and analytics.


# Test Performance:

## T-Test: Waiting Time vs. Readmission -
We perform the T-Test to check if average waiting time differs between readmitted and non-readmitted patients. A significant result would suggest waiting time impacts readmission rates, guiding reduction strategies.

```{r}
t_test_result <- t.test(`Waiting Time` ~ readmitted, data = data)
t_test_result
```
#### Insights: The T-test shows no significant difference in average waiting times between readmitted and non-readmitted patients (p = 0.5529, which is > 0.05). This suggests that waiting time is not a major factor influencing readmission rates.




## ANOVA: Waiting Time vs. Medical Specialty
We perform ANOVA to check if average waiting times significantly differ across different medical specialties. If significant, it helps identify departments causing longer waits, aiding in staff optimization.

```{r}
anova_result <- aov(`Waiting Time` ~ medical_specialty, data = data)
summary(anova_result)
```
#### Insights: The ANOVA test shows no significant difference in average waiting times across medical specialties (p = 0.311, which is > 0.05). This suggests that waiting time is not strongly influenced by the medical specialty.



## Chi-Square Test: Readmission vs. Ward Name
We perform the Chi-Square test to check if readmission rates vary across different hospital wards. A significant result would suggest that certain wards have higher readmission risks, helping in targeted interventions.

```{r}
table_ward <- table(data$readmitted, data$`Ward Name`)
chisq_test_ward <- chisq.test(table_ward)
chisq_test_ward
```
#### Insights:The Chi-Square test shows no significant association between ward name and readmission rates (p = 0.1193, which is > 0.05). This suggests that readmission likelihood does not vary significantly across different hospital wards.



#Logistic Regression:

```{r}
# Convert categorical variables to factors
data$age <- as.factor(data$age)
data$`Length of Stay` <- as.factor(data$`Length of Stay`)
data$medical_specialty <- as.factor(data$medical_specialty)
data$`Ward Name` <- as.factor(data$`Ward Name`)
data$`Time Visited Hospital` <- as.factor(data$`Time Visited Hospital`)
data$readmitted <- as.factor(data$readmitted)

# Select relevant columns
data <- dplyr::select(data, -`Patient ID`, -`Nurses Attending`, -`Doctors Attending`)

# Store the target variable before encoding
target <- data$readmitted  

# Store the target variable before encoding
target <- data$readmitted  

# Convert categorical variables using one-hot encoding (excluding readmitted)
data <- dplyr::select(data, -readmitted)  # Ensure proper selection
dummy_model <- dummyVars(" ~ .", data = data)
data <- predict(dummy_model, newdata = data) %>% as.data.frame()

# Restore the target variable
data$readmitted <- target
data$readmitted <- as.factor(data$readmitted)  # Ensure it's a factor

# Restore the target variable
data$readmitted <- target

# Ensure readmitted is still a factor
data$readmitted <- as.factor(data$readmitted)

# Split into training and testing sets
set.seed(123)
trainIndex <- createDataPartition(data$readmitted, p = 0.7, list = FALSE)
trainData <- data[trainIndex, ]
testData <- data[-trainIndex, ]

# Train logistic regression model
model <- glm(readmitted ~ ., data = trainData, family = binomial)

# Model summary
summary(model)

# Predictions on test data
testData$predicted <- predict(model, newdata = testData, type = "response")
testData$predicted_class <- ifelse(testData$predicted > 0.5, 1, 0)

# Model evaluation
conf_matrix <- table(testData$readmitted, testData$predicted_class)
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
print(paste("Model Accuracy:", round(accuracy, 4)*100))

```



```{r}
# Print available coefficient names to check transformations
print(rownames(summary(model)$coefficients))

# Recommendations
cat("\nRecommendations:")

# Identify transformed column names
waiting_time_col <- grep("Waiting Time", rownames(summary(model)$coefficients), value = TRUE)
glucose_test_col <- grep("glucose_test", rownames(summary(model)$coefficients), value = TRUE)

# Check if 'Waiting Time' variable is significant
if (length(waiting_time_col) > 0 && summary(model)$coefficients[waiting_time_col, "Pr(>|z|)"] < 0.05) {
  cat("\n- Reduce waiting time to decrease readmission rates.")
}

# Check if 'glucose_test' variable is significant
if (length(glucose_test_col) > 0 && summary(model)$coefficients[glucose_test_col, "Pr(>|z|)"] < 0.05) {
  cat("\n- Monitor glucose levels closely for better patient outcomes.")
}

```

