---
title: "Final_Report"
author: "Gauri Nagre"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(readxl)
library(tidyr)
library(stats)
library(caret)
library(ROCR)
library(MASS)  # For stepwise selection
library(randomForest)  # For improved accuracy
```


## Load Data
```{r}
dataset <- read_excel("C:/Users/Gauri Vijay Nagre/Desktop/Sem_2/BA/BA_Lifeline_Hospital_Dashboard/Hospital Dataset3.xlsx", sheet = "Sheet1")
head(dataset)
```


## Objectives:
#### 1. Reduce patient readmission rates by addressing underlying factors. 
#### 2. Optimize staff scheduling to align with peak patient visit times. 
#### 3. Shorten patient waiting times to improve satisfaction and care quality. 
#### 4. Enhance operational efficiency through targeted interventions and analytics.


# Test Performance:

## T-Test: Waiting Time vs. Readmission -
We perform the T-Test to check if average waiting time differs between readmitted and non-readmitted patients. A significant result would suggest waiting time impacts readmission rates, guiding reduction strategies.

```{r}
t_test_result <- t.test(`Waiting Time` ~ readmitted, data = dataset)
t_test_result
```
### Insights: The T-test shows no significant difference in average waiting times between readmitted and non-readmitted patients (p = 0.5529, which is > 0.05). This suggests that waiting time is not a major factor influencing readmission rates.




## ANOVA: Waiting Time vs. Medical Specialty
We perform ANOVA to check if average waiting times significantly differ across different medical specialties. If significant, it helps identify departments causing longer waits, aiding in staff optimization.

```{r}
anova_result <- aov(`Waiting Time` ~ medical_specialty, data = dataset)
summary(anova_result)
```
#### Insights: The ANOVA test shows no significant difference in average waiting times across medical specialties (p = 0.311, which is > 0.05). This suggests that waiting time is not strongly influenced by the medical specialty.



## Chi-Square Test: Readmission vs. Ward Name
We perform the Chi-Square test to check if readmission rates vary across different hospital wards. A significant result would suggest that certain wards have higher readmission risks, helping in targeted interventions.

```{r}
table_ward <- table(dataset$readmitted, dataset$`Ward Name`)
chisq_test_ward <- chisq.test(table_ward)
chisq_test_ward
```
#### Insights:The Chi-Square test shows no significant association between ward name and readmission rates (p = 0.1193, which is > 0.05). This suggests that readmission likelihood does not vary significantly across different hospital wards.




```{r}
names(dataset)
```



## Data Preprocessing
```{r}
# Convert categorical variables to factors
dataset <- dataset %>% mutate(across(where(is.character), as.factor))

# Remove Patient ID as it's not useful for prediction
dataset <- dataset %>% select(-contains("Patient ID"))

# Handle missing values
dataset <- na.omit(dataset)
```

## Train-Test Split
```{r}
set.seed(123)
trainIndex <- createDataPartition(dataset$readmitted, p = 0.7, list = FALSE)
trainData <- dataset[trainIndex, ]
testData <- dataset[-trainIndex, ]
```

## Logistic Regression Model
```{r}
log_model <- glm(readmitted ~ ., data = trainData, family = binomial)
summary(log_model)
```

## Predictions & Accuracy (Logistic Regression)
```{r}
pred_probs <- predict(log_model, testData, type = "response")
pred_labels <- ifelse(pred_probs > 0.5, 1, 0)
conf_matrix <- table(pred_labels, testData$readmitted)
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
paste("Logistic Regression Accuracy:", round(accuracy * 100, 2), "%")
```

## Random Forest Model
```{r}
set.seed(123)
rf_model <- randomForest(readmitted ~ ., data = trainData, ntree = 500, mtry = 3)
print(rf_model)
```

## Predictions & Accuracy (Random Forest)
```{r}
pred_rf <- predict(rf_model, testData)
conf_matrix_rf <- table(pred_rf, testData$readmitted)
accuracy_rf <- sum(diag(conf_matrix_rf)) / sum(conf_matrix_rf)
paste("Random Forest Accuracy:", round(accuracy_rf * 100, 2), "%")
```

## Recommendations
1. **Optimize Waiting Times**: Reduce patient waiting times to minimize readmissions.
2. **Focus on High-Risk Departments**: Allocate more resources to specialties with high readmission rates.
3. **Improve Discharge Planning**: Ensure follow-up care for patients at risk of readmission.
4. **Use Predictive Model Insights**: Apply the model to proactively identify at-risk patients and intervene early.
